% Calculus II Final (Original Feb. 28, 1995)

\documentstyle[12pt]{jarticle}

%A4 Size Setting
\topmargin = 0cm
\oddsidemargin = 0cm \evensidemargin = 0cm
\textheight = 23cm \textwidth = 15cm % default 16cm
%A4 size Setting End

\begin{document}
\begin{center}
{\gt\LARGE Calculus II  FINAL}\\
{\gt Winter 1995}
\end{center}

\noindent
ID 番号、氏名を、各解答用紙に、また、問題番号も忘れずに書いて下さい。\\
(Write your ID number and your name on each of your solution sheet. Do not forget to write the problem number as well.)

\begin{enumerate}
\item $(x,y)$ が、$(0,0)$ に近づくとき、次の関数の極限は存在するか、決定し、理由も述べよ。(Determine whether the limit of the following function exists or not when $(x,y)$ approaches to $(0,0)$. Give your reason.)
$$\frac{xy^2}{x^2+y^4}.$$

\item 次の関数の、極大値、極小値を決定せよ。(Determine maximum and minimum of the following function.)
$$xy(ax+by+c), \; (abc > 0).$$

\item 次の積分の順序を変更せよ。(Change the order of the integrals of the following.)
$$\int^2_0\left\{\int^{\sqrt{2x}}_{\sqrt{2x-x^2}}f(x,y)dy\right\}dx.$$

\item 次の重積分を求めよ。(Find the following multiple integrals.)
     \begin{enumerate}
     \item $\displaystyle{\int\!\int_D (x^2+y^2)dxdy, \quad D = \{(x,y)| 0\leq x, \:0\leq y, \:x+y \leq 1\}}.$
     \item $\displaystyle{\int\!\int_D xydxdy, \quad D = \{(x,y)| \frac{x^2}{a^2}+\frac{y^2}{b^2}\leq 1\}} \:(a >0, \:b>0).$
     \item $\displaystyle{\int\!\int_D \frac{y}{x+y}e^{(\frac{y}{x+y})^2}dxdy, \quad D = \{(x,y)| 0\leq x+y \leq 2,\: x\geq 0,\: y\geq 0\}}.$\\
     (Hint: $u=x+y,\:v=y/(x+y)$.)
     \item $\displaystyle{\int\!\int\!\int_V \sqrt{x^2+y^2+z^2}dxdydz, \quad V = \{(x,y,z)| x^2+y^2+z^2\leq 1\}}$
     \end{enumerate}

\item 次の級数の収束、発散を決定せよ。(Determine whether each of the following seires converges or diverges.)
     \begin{enumerate}
     \item $\displaystyle{\sum_{n=1}^{\infty}\frac{n+1}{n^2+1}}.$
     \item $\displaystyle{\sum_{n=1}^{\infty}\frac{1\cdot 3 \cdots (2n-1)}{2\cdot 5 \cdots (3n-1)}}.$
     \end{enumerate}

\item 次のべき級数の収束半径を求めよ。(Determine the radius of convergence of the following power series.)
     \begin{enumerate}
     \item $\displaystyle{\sum_{n=1}^{\infty}(\sqrt{n+1}-\sqrt{n})x^n}.$
     \item $\displaystyle{\sum_{n=1}^{\infty}\frac{(n!)^2}{(2n)!}x^n}.$
     \end{enumerate}

\vspace{3ex}
\noindent
次の２問（７、８）の中から、１問を選択し、解答せよ。(Choose one of the following problems (7 or 8) and solve it.)
     
\item 次を示せ。(Show the following.)
$$\log(1+x) = \sum_{n=1}^{\infty}(-1)^{n-1}\frac{x^n}{n} \quad(|x|<1).$$

\item 半球面 $x^2+y^2+z^2 = 1 \:(z\geq 0)$ の円柱 $x^2+y^2 = x$ の内部の部分の面積を求めよ。(Find the area of a half sphere $x^2+y^2+z^2 = 1 \:(z\geq 0)$ bounded by a cylinder $x^2+y^2 = x$. 
\end{enumerate}
     

\end{document}
