% Original Taken From Calculus II Final (Original Feb. 28, 1995)
% Calculus II TAKE-HOME MidTerm to be distributed on January 9, 1997

\documentstyle[12pt]{jarticle}

%A4 Size Setting
\topmargin = 0cm
\oddsidemargin = 0cm \evensidemargin = 0cm
\textheight = 23cm \textwidth = 16cm
%A4 size Setting End

\pagestyle{empty}

\begin{document}
\begin{center}
{\gt\LARGE Calculus II  Take-Home Midterm}\\
{\gt Winter Term, AY1996-7}
\end{center}

\bigskip
\noindent
解答は、ルーズリーフまたは、レポート用紙を使用し、ID 番号、氏名、また、問題番号も忘れずに書いて下さい。１月１３日（月）午後４時までに、鈴木研究室（ＮＳ-Ｓ３０９）前に用意した箱に提出して下さい。特別な理由で、この時間までに提出できない場合は、その旨、この時間以前に、鈴木に連絡すること。（電話、留守電も可。0422-33-3292）

\begin{enumerate}
\item $f(x,y)$ を以下のような関数とする。(Let $f(x,y)$ be a function defined as follows.)
$$f(x,y) = \left\{\begin{array}{cl} \frac{xy^3}{x^2+y^6} & (x,y)\neq (0,0)\\ 0 & (x,y) = (0,0)\end{array}\right..$$
	\begin{enumerate}
	\item $f(x,y)$ は、点 $(0,0)$ で連続かどうか判定し、理由ものべよ。
	(Determine whether the function $f(x,y)$ is continuous at $(0,0)$. Give your reason.)
	\item $f(x,y)$ は、点 $(0,0)$ で $x$ に関して偏微分可能かどうか判定し、理由ものべよ。
	(Determine whether the partial derivative of the function $f(x,y)$ with respect to $x$ exists at $(0,0)$. 
	Give your reason.)
	\item $f(x,y)$ は、点 $(0,0)$ で全微分可能かどうか判定し、理由ものべよ。
	(Determine whether the total derivative of the function $f(x,y)$ exists at $(0,0)$. 
	Give your reason.)
	\item $z = f(x,y)$ の点 $(2,1,f(2,1))$ での接平面の方程式を求めよ。
	(Find the equation of the tangent plane of the surface $z = f(x,y)$ at $(2,1,f(2,1)$.)
	\end{enumerate}

\item $z = \cos(x/y)$ とする。(Let $z = \cos(x/y)$.)
     \begin{enumerate}
     \item $x$ と、$y$ に関する $z$ の偏導関数を求めよ。(Find the partial derivatives of $z$ with respect to $x$ and $y$.)
     \item $z$ は、次の式を満たすことを示せ。(Show that $z$ satisfies the following equation.)
     $$x^2\frac{\partial^2z}{\partial x^2} + 2xy\frac{\partial^2z}{\partial y\partial x} + y^2\frac{\partial^2z}{\partial y^2} = 0.$$
     \end{enumerate}

\item 次の式で与えられた合成関数について、$u$, $v$ に関する偏導関数を求めよ。(Find the partial derivatives of the following composite functions with respect to $u$ and $v$.)
$$z = e^{xy}, \;x = \sin uv, \;y = \cos(u+v)$$
 
\item 次の関数の、極大値、極小値を決定せよ。(Determine maximum and minimum of the following function.)
     \begin{enumerate}
     \item $f(x,y) = x^3 + 3xy - y^3$。
     \item $f(x,y) = xy(ax+by+c), \; (abc > 0)$。
     \end{enumerate}
\end{enumerate}
     

\end{document}
