% Original Taken From Calculus II Final (Original Feb. 28, 1995)
% Calculus II TAKE-HOME MidTerm to be distributed on January 9, 1997
% Calculus II Final, on February 28, 1997, at 8:30
%% original: February 26, 27, 1997.

\documentstyle[12pt]{jarticle}

%A4 Size Setting
\topmargin = -0.5cm
\oddsidemargin = 0cm \evensidemargin = 0cm
\textheight = 23.5cm \textwidth = 16cm
%A4 size Setting End

\pagestyle{empty}

\newtheorem{thm}{定理}
\newcommand{\proof}{{\gt 証明\quad}}
\newcommand{\qed}{\hfill\hbox{\rule{6pt}{6pt}}}

\def\inkern{\mathchoice{\!\!\!}{\!\!}{\!\!}{\!\!}}
\def\iint{\int\inkern\int} %double iterated integral
\def\iiint{\int\inkern\int\inkern\int} %triple iterated integral

\begin{document}
\begin{center}
{\gt\LARGE Calculus II  Final}\\
{\gt Winter Term, AY1996-7}
\end{center}

\bigskip
\noindent
ID 番号、氏名を、各解答用紙に、また、問題番号も忘れずに書いて下さい。\\
(Write your ID number and your name on each of your solution sheet. Do not forget to write the problem number as well.)

\begin{enumerate}
\item 曲面、${\displaystyle z = \frac{x^2}{3^2} + \frac{y^2}{5^2}}$ の、点 $(3,-5,2)$ における接平面の方程式を求めよ。(Find the equation of the plane tangent to a surface 
${\displaystyle z = \frac{x^2}{3^2} + \frac{y^2}{5^2}}$ at the point $(3,-5,2)$.)

\item 原点から、曲面 $z^2 = xy - 5x - 4y + 16$ 上の点までの距離の最小値を求めよ。
(Determine the minimal distance between the origin and the points on the surface 
$z^2 = xy - 5x - 4y + 16$.)

\item 下の定理は、ラグランジュの乗数定理と呼ばれる。この証明の、各部分の空欄を埋める式を解答用紙に記入せよ。(The following is called Lagrange's multiplier theorem. Fill the empty boxes with the suitable formulas in the proof.)
\begin{thm}
$g(x,y) = 0$ の条件のもとで、関数 $f(x,y)$ が、点 $(a,b)$ で極値を持つとする。ただし、$g_x(a,b)$ と、$g_y(a,b)$ は、同時には、$0$ にならないものとする。そのとき、
$$f_x(a,b) + \lambda g_x(a,b) = 0,\quad f_y(a,b) + \lambda g_y(a,b) = 0$$
を満たす定数 $\lambda$ が存在する。
\end{thm}
\proof
$g_y(x,y) \neq 0$ とする。$g(x,y) = 0$、$g_y(x,y) \neq 0$ より、陰関数の定理により、$y = \phi(x)$ とかける。$g(x,\phi(x)) = 0$  を合成関数として、$x$ で微分すると、

$(a)$ \framebox[10cm][l]{\rule[-0.3cm]{0cm}{0.7cm}}

が得られる。従って、

$(b)$ \framebox[10cm][l]{\rule[-0.3cm]{0cm}{0.7cm}${\displaystyle \frac{d\phi}{dx} = }$}

一方、$f(x,\phi(x))$ が、$(a,b)$ で極値を取るという条件から、

$(c)$ \framebox[10cm][l]{\rule[-0.3cm]{0cm}{0.7cm}}

である。この式に、上で求めた、$\phi'(a)$  を代入して整理すると次の式が得られる。

$(d)$ \framebox[10cm][l]{\rule[-0.3cm]{0cm}{0.7cm}}

この比の値を $-\lambda$ と置くと、定理の結果を得る。

$g_x(x,y) \neq 0$ の時も同様。
\qed

\item 次の積分の順序を変更せよ。ただし、$a>0$ とし、答えは、いくつかの積分の和になっても構わない。(Change the order of the integrals of the following.  Here $a>0$ and the solution may be a sum of several integrals.)
$$\int^a_0\left\{\int^{2a - x}_{x^2/a}f(x,y)dy\right\}dx.$$

\item 下の変数変換を用いて、重積分を求めよ。(Evaluate the integral by changing the variables as indicated below.)
$$\iint_D \frac{x^2+y^2}{(x+y)^3}dxdy,\quad D = \{(x,y)\mid 1\leq x+y\leq 2, x\geq 0,y\geq 0\}, \:(x + y = u, y = uv).$$


\item 次の重積分を求めよ。(Evaluate the following multiple integrals.)
$$\iint_D \sin(x+y)dxdy, \quad D = \{(x,y)| 0\leq x, \:0\leq y, \:x+y \leq\frac{\pi}2\}.$$

\item 球 $x^2 + y^2 + z^2 \leq a^2 \:(a>0)$ の円柱 $x^2 + y^2 = ax$ の内部にある部分の体積を求めよ。(Find the volume of a part of a sphere $x^2 + y^2 + z^2 \leq a^2 \:(a>0)$ bounded by a cylinder $x^2 + y^2 = ax$.)

\item 次の級数の収束、発散を決定せよ。(Determine whether each of the following seires converges or diverges.)
     \begin{enumerate}
     \item $\displaystyle{\sum_{n=1}^{\infty}\Bigl(\sin\frac1n\Bigr)^n}.$
     \item $\displaystyle{\sum_{n=1}^{\infty}(\sqrt{n^2 + 1} - n)}.$
     \end{enumerate}

\item 次のべき級数の収束半径を求めよ。(Determine the radius of convergence of the following power series.)
     \begin{enumerate}
     \item $\alpha$ を実数としたとき、
$$1 + \alpha x + \frac{\alpha(\alpha-1)}{2}x^2 + \cdots + \frac{\alpha(\alpha-1)\cdots(\alpha - n + 1)}{n!}x^n + \cdots. $$
     \item $\displaystyle{\sum_{n=1}^{\infty}(2^n + n)x^n}.$
     \end{enumerate}

\vspace{3ex}
\noindent
次の２問の中から、１問を選択し、解答せよ。(Choose one of the following problems and solve it.)
     
\item ${\displaystyle \frac{1}{1-3x + 2x^2}}$ をべき級数で表し、その収束半径を求めよ。(Express ${\displaystyle \frac{1}{1-3x + 2x^2}}$ as a power series and find the radius of convergence.)

\item 曲面 $z = xy$ の 円柱 $x^2 + y^2 \leq a^2$  の内部の部分の面積を求めよ。(Find the area of a surface $z = xy$ bounded by a cylinder $x^2 + y^2 \leq a^2$. )
\end{enumerate}
    

\end{document}
%%%%%%%%%%%%%%%%%%%% End of Document %%%%%%%%%%%%

\item $f(x,y) = 4xy - 2y^2 - x^4$ とする。
\begin{enumerate}
\item 停留点をすべて求めよ。(Find all stationary points.)
\item 停留点が極点かどうかを判定し、極点の場合には、極値を求めよ。(Determine extremum and find the values at each relative maximum and relative minimum point.)
\end{enumerate}

\item 極座標を用いて次の重積分の値を求めよ。(Find the value of the following integral using polar coordinates.)
$$\iint_D\sqrt{x^2+y^2}dxdy,\quad D = \{(x,y)\mid x^2 + y^2 \leq y\}$$
